---
title: "Day 4 - Lecture 7 & 8"
author: "Ruth Enriquez"
date: "2022-08-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lecture 7

```{r}
#a nested for loop example

file_prefix <- c("temp", "ph", "salinity")
file_suffix <- c(1, 2, 3, 4, 5)

for (i in seq_along(file_prefix)) {
  for (j in seq_along(file_suffix)) {
    print(paste0(file_prefix[i], "_", file_suffix[j]))
  }
}
```

Slide: you try! what would be printed from...

```{r}
odds <- c(1, 3, 5)
evens <-c(2, 4, 6, 8)

for (i in seq_along(odds)) {
  for (j in seq_along(evens)) {
    print (odds[i] * evens[j])
  }
}
```

functions: self-contained bits of code that accept user input(s), do something with it, and return a useful output.

you should consider writing a function whenever you've copied and pasted a block of code more than twice (i.e. you now have three copies of the same code)

benefits of writing functions

1.  you can give it a name and it can make things easier, give it a name to help yourself out
2.  as requirements change, you only need to update code in one place, instead of many. you can just call it from where it was written

Anatomy of a function - R

```{r}
function_name <- function(bird, dog) {
  pets = bird + dog
  return(pets)
}
```

A function in R: that exclaims age input

```{r}
exclaim_age <- function(age){
  print(paste("I am", age, "years old!"))
}

exclaim_age(age = 10)
```

the name of a function is important. Ideally, the name of your function will be short, but clearly evoke what the function does. That's hard! But it's better to be clear than short, as RStudio's autocomplete makes it easy to type long names.

write a loop that returns your total time for any two consecutive quarter miles

```{r}
quarter_splits <-c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

for (i in seq_along(quarter_splits)) {
  half_time <- quarter_splits[i] + quarter_splits[i + 1]
  print(half_time)
}

```

A function with conditions: we can combine what we've already learned when building new functions.

What does this R function do?

```{r}
animal_age2 <- function(animal, age){
  if (animal == "dog") {
    print(age * 7)
  } else if (animal == "goat") {
    print(age * 4.7)
  }
}
```

Intro to functions double whatever value a user gives it

```{r}
double_this <- function(my_value) {
  print(2 * my_value)
}

double_this(my_value = 200)
double_this(200)

double_this(c(1, 2, 10, 100, 175))
```

Your turn: Write a function that ask for a user's NAME and age, and returns an exclamation, like "I'm [NAME], and I am [age] years old."

```{r}
exclaim_age <- function(name, age){
  print(paste("My name is", name, "and I am", age, "years old"))
}

exclaim_age(name = "Ruth", age = 27)
```

```{r}
convert_animal_age <- function(animal, human_age) {
  if (animal == "dog") {
    print(human_age * 7)
  } else if (animal == "goat") {
    print(human_age * 4.7)
  } else {
    print("Please enter dog or goat for animal")
  }
}

convert_animal_age(animal = "dog", human_age = 12)
convert_animal_age(animal = "goat", human_age = 27)
convert_animal_age(animal = "whale", human_age = 28)
```

Make a dataframe

```{r}
dog_choice <- tribble(
  ~dog_name, ~food,
  "Khora", "everything",
  "Teddy", "salmon",
  "Waffle", "pancakes",
  "Banjo", "chicken"
)

dog_menu <- function(name) {
  my_sub <- filter(dog_choice, dog_name == name)
  print(paste("My name is", my_sub$dog_name, "and my favorite food is", my_sub$food))
}

dog_menu(name = "Waffle")

```

Functions meet for loops

```{r}
means_over_columns <- function(df) {
  for (i in seq_along(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    column_name <- colnames(df[i])
    print(paste("The mean value of", column_name, "is", col_mean))
  }
}

means_over_columns(df = mtcars)
means_over_columns(df = faithful)
```

Comprehension check:

```{r}
apples <- c(1, 2, 3, 4)
price <- c(7, 10, 25)

fruit_out <- matrix(nrow = length(price), ncol = length(apples))

for (i in seq_along(apples)) {
  for (j in seq_along(price)) {
    total_cost <- price[j] * apples[i]
    fruit_out [j, i] <- total_cost
  }
}

#fruit_out [j, i] gives you the location in the matrix that will be populated
```

Nested for loop example

```{r}
coffee <- c(1, 2, 3, 4, 5)
price <- c(3.50, 4.50, 5.00)

coffee_price <- matrix(nrow = length(price), ncol = length(coffee))

for (i in seq_along(coffee)) {
  for (j in seq_along(price)) {
    total_cost <- price[j] * coffee[i]
    coffee_price [j, i] <- total_cost
  }
}
```

```{r}
clothing <- c("shirt", "pant", "shoes")
color <- c("purple", "red", "blue", "white")

our_clothes <- matrix(nrow = length(color), ncol = length(clothing))

for (i in seq_along(clothing)) {
  for (j in seq_along(color)) {
    clothing_details <- paste(color[j], clothing[i])
    our_clothes[j,i] <-clothing_details
  }
}

our_clothes
```

# Lecture 8

Starting point: What do I want to do?

I want to automate estimating the population size over a range of times, and at multiple growth rates (for a constant initial population and carrying capacity).

What might a matrix of population size outputs look like for different times and growth rates?

Function for logistic growth make sure that it works first

```{r}
logistic_growth <- function(N0, K, r, time) {
  Nt <- K / (1 + ((K - N0) /N0) * exp(-r * time))
  print(Nt)
}
```

FROM the slides: #\`\`\`{r} remove \# here to bring it to life #Create a sequence of growth rate values: time_vec \<- seq(from = 0, to = 35, by = 0.1) r_seq \<- seq(from = 0.20, to = 0.40, by = 0.01)

# Need to create a MATRIX to store the outputs in:

out_matrix \<- matrix(nrow = length(time_vec), ncol = length(r_seq))

#Now, a nested for loop:

for (i in 1:length(r_seq)) {# Outer loop is the growth rates for (j in 1:length(time_vec)) { \# Inner loop is the time sequence values pop \<- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j]) out_matrix[j, i] \<-pop #store the value in the appropriate row & column } } #\`\`\` remove \# to bring it to life

Logistic growth

Modeling a population over time and varying growth rate

```{r}
my_function <- function(oranges, rainbows) {
  (oranges * rainbows / 2.4)^2
}

# Shortcut: Mac: Cmd + Option + x
# PC: Ctrl + Alt + x
```

```{r}
logistic_growth <- function(K, N0, r, time) {
  K / (1 + ((K - N0) / N0) * exp(-r * time))
}

#Test:
logistic_growth(K = 5000, N = 100, r = 0.07, time = 42)
```

```{r}
time_vec <- seq(from = 0, to = 35, by = 0.1)

pop_35 <- logistic_growth(N0 = 40, K = 8000, r = 0.7, time = time_vec) #it's ok to be out of order IF you are explicit about what the # ='s to. write out the names to be safe

pop_time35_df <- data.frame(time_vec, pop_35)# vectors need to be the same length
```

```{r}
#Make a ggplot

ggplot(data = pop_time35_df, aes(x = time_vec, y = pop_35))+
  geom_line()
```

Put this into a for loop

```{r}
pop_35_vec <-vector(mode = "numeric", length = length(time_vec))

for (i in seq_along(time_vec)) {
  population <- logistic_growth(N0 = 100, K = 6000, r = 0.27, time_vec[i])
  pop_35_vec[i] <- population
}
```

Now extend this to iterate over multiple growth rates:

```{r}
r_seq <- seq(from = 0.20, to = 0.4, by = 0.01)
```

```{r}
out_matrix <-matrix(nrow = length(time_vec), ncol = length(r_seq))

for (i in seq_along(r_seq)){
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(N0 = 100, K = 6000, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}

### Now some wrangling
logistic_df <- data.frame(out_matrix, time = time_vec)

colnames(logistic_df) <- c(paste0("gr_", r_seq), "time")

#we have three variables here: time, population, growth rate
#PIVOT
# -time, use ALL the columns but EXCLUDE time column

logistic_df_long <- logistic_df |> 
  pivot_longer(cols = -time, values_to = "population", names_to = "growth_rate")

ggplot(data = logistic_df_long, aes(x = time, y = population))+
  geom_line(aes(color = growth_rate))
```

## A function with warnings and an error message

Use `warning` to add a warning.

Use `stop` to add an error.



$$
P=0.3\rho A v^3
$$
rho: air density (kg / m^3)
A: area covered by turbine (square meters)
v: wind velocity (m/s)

```{r}
calc_windpower <- function(rho, radius, windspeed) {
  
  if(windspeed > 130){
    warning("whoa, that's really fast. Are you sure?")
  }
  if (rho > 1.225){
    warning("That air density is suspicious")
  }
  
  if (radius < 0) {
    stop("Rotor radius must be a positive value (meters).")
  }
  0.3*rho*pi*(radius^2)*(windspeed^3)
}

```

